Storm Surges

### 2014-07-08 RoughWork/manuscript.tex
Added notes to include spatial extent of surge

\subsubsection{Tidal forcing} 
The model was forced by eight tidal constituents ($K_1$ ,$O_1$, $P_1$, $Q_1$, $M_2$, $K_2$, $N_2$, $S_2$) at the Juan de Fuca and Johnstone Strait boundaries. Tidal heights and currents at grid points along the Juan de Fuca boundary were extracted from Webtide, an online web prediction model for the northeast Pacific Ocean, which is based on work by \citet{foreman2000webtide}. The Johnstone Strait boundary was forced with current and elevation tidal harmonics measured and calculated by \citet{thomson1980johnstone} for the major $M_2$ and $K_1$ constituents. Additionally, $O_1$ and $S_2$ elevation harmonics from their measurements were employed. The remaining constituents were extrapolated from Webtide.

@@

\subsection{Tidal evaluation}
The model was initially evaluated qualitatively by comparing patterns of tidal amplitude and phase to results from \citep{foreman1995tidal}. For example, the amphidromic dome around Victoria was produced in the $M_2$ results, as well as the monotonic increase in $K_1$ amplitude moving northwards along the Strait of Georgia. Initially, the Johnstone Strait boundary was closed, but modelled $M_2$ amplitudes were too small compared to measured amplitudes... TBC

@@

Once our model was reproducing observed tidal patterns, model results were quantitatively evaluated by comparing modelled harmonic constituents to measured harmonic constituents at tidal measuring stations throughout the domain. Comparisons were made using the complex difference (D), defined by \citep{foreman1995tidal} as:

@@

\begin{table}[h]
\centering 
\begin{tabular}{l c c c c c c} 
\hline 
& \multicolumn{3}{c}{$M_2$} & \multicolumn{3}{c}{$K_1$} \\ 
\hline 
Location   & $R$ & $ \Delta \phi$ (deg) & $D$ (cm) & $R$ & $ \Delta \phi$ (deg) & $D$ (cm) \\
\hline 
           &     &                      &          &     &                      &          \\
\hline   
Mean Error &     &                      &          &     &                      &          \\
RMS Error  &     &                      &          &     &                      &          \\
\hline  
\end{tabular}
\caption{Comparisons of modelled and observed amplitudes and phases.}
\label{tab:comparison} 
\end{table}
%May have to include station numbers and locations on a map? Might be too busy with rivers too...
%Should we use the new method of calculating tides?

@@

Three significiant surge or wind events have been considered. First, a study of a large surge event on Feb 4, 2006 is presented. This event caused siginificant damage and flooding in Boundary Bay, a community south of Richmond. (citation) and has been chosen as a case study to demonstrate the model's skill in reproducing a storm surge and to quantitatively assess forcing conditions that are important in storm surge hindcasting. 
%% Trying to decide Dec 15, 2005 or Nov 2009. I think it's more sense to do Nov 2009 - strong winds but no surge. 
Next, an extreme wind event which resulted in damage to property in Vancouver and surrounding areas is discussed. This event occured on Dec 15, 2006 but did not result in significant flooding such as that which occured on Feb 4, 2006 because the storm did not occur at high tide. This case was choosen to idenitify the model's capability in handling large wind events that do not result in high water levels. Finally, a case with high resolution atmospheric forcing for a recent surge on Dec 17, 2012 is discussed.

\subsection{Factors Contributing to Storm Surges: Feb 4, 2006}

@@

The model's ability to reproduce this event is demonstrated in Figure \ref{fig:feb2006} which compares observations and corrected model output at Point Atkinson, Victoria, Patricia Bay, and Campbell River. The timing of maximum water level matches well between the observations and the model at all four locations, the most siginificant contribution to high water level being from the tides. There is also agood match in amplitude of the water level maximum at Point Atkinson and Cambell River, with a relative error less than 
2\%. However, the error at Victoria and Patricia Bay is more siginificant, at 12\% and 8\% respectively, likely due to error in $M_2$ tidal amplitudes as discussed above. The precise location and maginitude of the $M_2$ amphidromic dome around Victoria is very sensitive to model parameters such as bottom friction.

The error in modelled residuals at Victoria and Campbell River is also apparent in these figures. Most noticebale is underestimate in modelled residual throughout most of the time period. Additionally, there is a delay of about 2-3 hours in the timing of maximum residual at most locations and the  relative errors in maximum residual range from 5\% at Point Atkinson to 29\% at Patricia Bay. 
%Question: is it better to present relative errors or just difference?
%I think Hal Ritchie's talk with the storm surges had some stats presented. Look this up...

@@

\subsection{Strong wind event: Dec 15, 2006}

@@

\section{Acknowledgements}\label{sec:ack}
We would like to thank Michael Foreman, Diane Masson, and Luc Fillion for providing data used in model set up and evaluation. This project is funded by the Marine Environment Observation Prediction and Response Network (MEOPAR), a Network of Centres of Excellence of Canada.  
%I'm likely missing a lot more (Keith, Youyu, Vasily, JP, Rich, Mark). Not sure about authorship but something to think about for later....

>>>>

\subsubsection{Tidal forcing} 
The model was forced by eight tidal constituents (K$_1$ ,O$_1$, P$_1$, Q$_1$, M$_2$, K$_2$, N$_2$, S$_2$) at the Juan de Fuca and Johnstone Strait boundaries. Tidal heights and currents at grid points along the Juan de Fuca boundary were extracted from Webtide, an online web prediction model for the northeast Pacific Ocean, which is based on work by \citet{foreman2000webtide}. The Johnstone Strait boundary was forced with current and elevation tidal harmonics measured and calculated by \citet{thomson1980johnstone} for the major M$_2$ and K$_1$ constituents. Additionally, O$_1$ and S$_2$ elevation harmonics from their measurements were employed. The remaining constituents were extrapolated from Webtide.

@@

\subsection{Tidal evaluation}
The model was initially evaluated qualitatively by comparing patterns of tidal amplitude and phase to results from \citep{foreman1995tidal}. For example, the amphidromic dome around Victoria was produced in the M$_2$ results, as well as the monotonic increase in $K_1$ amplitude moving northwards along the Strait of Georgia. Initially, the Johnstone Strait boundary was closed, but modelled M$_2$ amplitudes were too small compared to measured amplitudes... TBC

@@

Once our model was reproducing observed tidal patterns, model results were quantitatively evaluated by comparing modelled harmonic constituents to measured harmonic constituents at tidal measuring stations throughout the domain. Comparisons were made using the complex difference ($D$), defined by \citep{foreman1995tidal} as:

@@

\begin{table}[h]
\centering 
\begin{tabular}{l c c c c c c} 
\hline 
& \multicolumn{3}{c}{M$_2$} & \multicolumn{3}{c}{K$_1$} \\ 
\hline 
Location   & $R$ & $ \Delta \phi$ (deg) & $D$ (cm) & $R$ & $ \Delta \phi$ (deg) & $D$ (cm) \\
\hline 
           &     &                      &          &     &                      &          \\
\hline   
Mean Error &     &                      &          &     &                      &          \\
RMS Error  &     &                      &          &     &                      &          \\
\hline  
\end{tabular}
\caption{Comparisons of modelled and observed amplitudes and phases. $R$ is the ration of modelled to observed amplitude. $\Delta \phi$ is the difference between modelled and observed phase. $D$ is the complex difference. }
\label{tab:comparison} 
\end{table}
%May have to include station numbers and locations on a map? Might be too busy with rivers too...
%Should we use the new method of calculating tides?

@@

Three significiant surge or wind events have been considered. First, a study of a large surge event on Feb 4, 2006 is presented. This event caused siginificant damage and flooding in Boundary Bay, a community south of Richmond. (citation) and has been chosen as a case study to demonstrate the model's skill in reproducing a storm surge and to quantitatively assess forcing conditions that are important in storm surge hindcasting. 
%% Trying to decide Dec 15, 2005 or Nov 2009. I think it's more sense to do Nov 2009 - strong winds but no surge. 
Next, an extreme wind event which resulted in property damage in Vancouver and surrounding areas is discussed. This event occured on Dec 15, 2006 but did not result in significant flooding such as that which occured on Feb 4, 2006 because the storm did not occur at high tide. This case was choosen to identify the model's capability in handling large wind events that do not result in high water levels. Finally, a case with high resolution atmospheric forcing for a recent surge on Dec 17, 2012 is discussed.

\subsection{Factors Contributing to Storm Surges: Feb 4, 2006}
%Include spatial extend of surge

@@

The model's ability to reproduce this event is demonstrated in Figure \ref{fig:feb2006} which compares observations and corrected model output at Point Atkinson, Victoria, Patricia Bay, and Campbell River. The timing of maximum water level matches well between the observations and the model at all four locations, the most siginificant contribution to high water level being from the tides. There is also a good match in amplitude of the water level maximum at Point Atkinson and Cambell River, with a relative error less than 2\%. However, the error at Victoria and Patricia Bay is more siginificant, at 12\% and 8\% respectively, likely due to error in M$_2$ tidal amplitudes as discussed above. The precise location and magnitude of the M$_2$ amphidrome around Victoria is very sensitive to model parameters such as bottom friction.

The error in modelled residuals at Victoria and Campbell River is also apparent in these figures. Most noticebale is underestimate in modelled residual throughout most of the time period. Additionally, there is a delay of about 2-3 hours in the timing of maximum residual and the relative errors in maximum residual range from 5\% at Point Atkinson to 29\% at Patricia Bay. 
%Question: is it better to present relative errors or just difference?
%I think Hal Ritchie's talk with the storm surges had some stats presented. Look this up...

@@

\subsection{Strong wind event: Dec 15, 2006}% or Nov 2009

@@

\section{Acknowledgements}\label{sec:ack}
We would like to thank Michael Foreman, Diane Masson, and Luc Fillion for providing data used in model set up and evaluation. This project is funded by the Marine Environmental Observation Prediction and Response Network (MEOPAR), a Network of Centres of Excellence of Canada.  
%I'm likely missing a lot more (Keith, Youyu, Vasily, JP, Rich, Mark). Not sure about authorship but something to think about for later....
